Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-11-18 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
started running lime for different numbers of bins

# Load libraries
library(tidyverse)
library(bulletr)
library(randomForest)
library(lime)
library(cowplot)

@@

```{r}
# Apply the lime function if the file does not already exist
if(!file.exists("../data/hamby224_lime.rds")) {

  # Apply the lime function from the lime package (with a seed set)
  # Note that the as_classifier must be added since rtrees is from the
  # randomForest package and not fit using caret or one of the other 
  # available models specified in the lime package. Additionally, the
  # randomForest package must be loaded in order to run the functions
  # from the lime package. (I should check on exactly how this works.)
  set.seed(84902)
  hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                        model = as_classifier(rtrees),
                        n_bins = 4)

  # Save the lime object
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")

@@

  # Load in the lime object
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")

@@

```{r}
# Apply explain function if the file does not already exist
if(!file.exists("../data/hamby224_explain.rds")) {

  # Apply the explain function from the lime package (with a seed set)
  set.seed(84902)
  hamby224_explain <- lime::explain(hamby224_test %>% 
                                      arrange(case) %>% 
                                      select(rf_features) %>% 
                                      na.omit(), 
                                    hamby224_lime,
                                    labels = "TRUE",
                                    n_features = 3)

  # Save the explainer object
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

@@

  # Load in the explain object
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

```{r}
# Create a dataframe with the bins
if (!file.exists("../data/hamby_bins.csv")) {

  # Determine the nubmer of bins
  nbins <- length(hamby224_lime$bin_cuts[[1]]) - 1

  # Create a dataframe of bin boundaries
  hamby_divisions <- data.frame(Feature = rf_features, 
                                Lower = c(0, -1, 0, 0, 0, 0, 0, 0, 0),
                           matrix(unlist(hamby224_lime$bin_cuts), nrow = 9, 
                                  byrow = TRUE)[,2:nbins],
                           Upper = c(1, 1, rep(NA, 7)))

  # Use my function "write_bins" to create a dataframe with the bins
  hamby_bins <- data.frame(t(apply(hamby_divisions, 1, write_bins)))

  # Assign appropriate names to the bin columns
  if (nbins == 2){
    names(hamby_bins) <- c("Feature", "Lower Bin", "Upper Bin")
  } else {
    names(hamby_bins) <- c("Feature", "Lower Bin", 
                           sapply(1:(nbins-2), 
                                  function(bin) sprintf("Middle Bin %.0f", bin)),
                           "Upper Bin")
  }

  # Save the bin cuts
  write.csv(hamby_bins, "../data/hamby_bins.csv", row.names = FALSE)

  # Load in the bins csv
  hamby_bins <- read.csv("../data/hamby_bins.csv")

  # Replace the periods with spaces in the feature names
  names(hamby_bins) <- gsub(".", " ", names(hamby_bins), fixed = TRUE)

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = mapply(bin_labeller, feature = feature, 
                                feature_value = feature_value)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE, 
           feature = factor(feature))

>>>>

# Load libraries
library(tidyverse)
library(bulletr)
library(randomForest)
library(lime)
library(cowplot)
library(purrr)

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../data/hamby224_lime.rds")) {

  # Specify the number of bins to use
  nbins = 2:10

  # Run the lime functions for the specified number of bins 
  hamby224_all_results <- map(.x = nbins, 
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3)

  # Separate the lime and explain function results
  hamby224_lime <- map(hamby224_all_results, function(list) list$lime)
  hamby224_explain <- map_df(hamby224_all_results, function(list) list$explain)

  # Save the lime and explain results objects
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

@@

  # Load in the lime results object
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../data/hamby224_bins.csv")) {

  # Create a list a dataframes with the bin boundaries and bins 
  # for the different evaluations of the lime functions
  hamby224_bin_list <- map(hamby224_lime, create_bin_data)

  # Save the bin boundaries and the bins as separate dataframes
  hamby224_bin_boundaries <- map(hamby224_bin_list, function(m) m$boundaries)
  hamby224_bins <- map(hamby224_bin_list, function(m) m$bins)

  # Save the bin boundaries and bins
  saveRDS(hamby224_bin_boundaries, "../data/hamby224_bin_boundaries.rds")
  saveRDS(hamby224_bins, "../data/hamby224_bins.rds")

@@

  # Load in the bin boundaries and bins
  hamby224_bin_boundaries <- readRDS("../data/hamby224_bin_boundaries.rds")
  hamby224_bins <- readRDS("../data/hamby224_bins.rds")

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain_results, by = "case") %>%
    mutate(case = factor(case),
           feature = factor(feature),
           feature_desc = factor(feature_desc)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

  # mutate(feature_bin = mapply(bin_labeller, feature = feature, 
  #                               feature_value = feature_value))
